### 2013-03-15 bwamem.tex
put figures in the base directory, not in eval/

\documentclass{bioinfo}
\copyrightyear{2013}
\pubyear{2013}

\usepackage{natbib}
\bibliographystyle{apalike}

\begin{document}
\firstpage{1}

\title{Aligning sequence reads, clone sequences and assembly contigs with BWA-MEM}

\author[Li]{Heng Li}

\address{Broad Institute of Harvard and MIT, 7 Cambridge Center, Cambridge, MA 02142, USA}

\history{Received on XXXXX; revised on XXXXX; accepted on XXXXX}
\editor{Associate Editor: XXXXXXX}
\maketitle

\begin{abstract}
\section{Summary:} BWA-MEM is a new alignment algorithm for aligning sequence
reads or long query sequences against a large reference genome such as human.
It automatically chooses between local and end-to-end alignments, supports
paired-end reads and performs split alignment. The algorithm is robust to
sequencing errors and applicable to a wide range of sequence lengths from 70bp
to a few megabases. For short-read mapping, BWA-MEM shows better performance
than several state-of-art read aligners to date.

\section{Availability and implementation:} BWA-MEM is implemented as a
component of BWA, which is available at http://github.com/lh3/bwa.

\section{Contact:} hengli@broadinstitute.org
\end{abstract}

\section{Introduction}

Most short-reads mappers for next-generation sequencing (NGS) data were
developed when sequence reads were about 36bp in length. For short reads, it is
reasonable to require end-to-end alignment (i.e. every read base to be aligned
to the reference) and to only report hits within certain hamming or edit
distance.  However, with emerging technologies and improved chemistry, NGS
reads are not short any more, which poses new challenges to read alignment. For
longer reads, it becomes more important to allow long gaps under the affine-gap
penalty and to report multiple non-overlapping local hits potentially caused by
structural variations or misassemblies in the reference genome. Many short-read
alignment algorithms are not applicable or not preferred for mapping longer
reads. At the same time, although several mature algorithms exist for aligning
capillary sequence reads, they are slow and lack features for analyzing
large-scale NGS data. Fast moving NGS technologies keep pressing for the
development of new alignment algorithms.

In this background, a few long-read alignment algorithms, including
BWA-SW~\citep{Li:2010fk}, Bowtie2~\citep{Langmead:2012fk},
Cushaw2~\citep{Liu:2012uq} and GEM~\citep{Marco-Sola:2012kx}, have been
developed. However, they all have some weakness. BWA-SW is slower than bowtie2
at a comparable accuracy and less accurate then Cushaw2 at a comparable speed
(see Results). Bowtie2 and Cushaw2 do not work with very long query sequences.
These three aligners do not fully use the paired-end information to improve
alignments in hard regions, either.  While GEM is both fast and accurate, it
mandates end-to-end alignment and does not perform affine-gap alignment, which
may limit its uses for long-read alignment. A fast, accurate and fully featured
aligner accepting sequences with a wide range of lengths is still lacking. This
motivates us to explore a new algorithm.

\begin{methods}

\section{Methods}

\subsection{Aligning a single query sequence}

\subsubsection{Seeding and re-seeding} BWA-MEM follows the canonical
seed-and-extend paradigm. It initially seeds an alignment with supermaximal
exact matches (SMEMs) using an algorithm we found previously~\citep[Algorithm
5]{Li:2012fk}, which essentially finds at each query position the longest exact
match covering the position. However, occasionally the true alignment may not
contain any SMEMs.  To reduce mismappings caused by missing seeds, we introduce
re-seeding. Suppose we have a SMEM of length $l$ with $k$ occurrences in the
reference genome. If $l$ is too large (over 28bp by default), we re-seed
with the longest exact matches that cover the middle base of the SMEM and occur
at least $k+1$ times in the genome. Such seeds can be found by requiring a
minimum occurrence in the original SMEM algorithm.

\subsubsection{Chaining and chain filtering} We call a group of seeds that are
colinear and close to each other as a \emph{chain}. We chain the seeds while
seeding. After chains are collected, we filter out short chains that are
largely contained in a long chain and are much worse than the long chain (by
default, both 50\% and 38bp shorter than the long chain). Chain filtering aims
to reduce unsuccessful seed extension at a later step. Each chain may not
always correspond to a final hit. Chains detected here do not need to be
accurate.

\subsubsection{Seed extension} We rank a seed by the length of the chain it
belongs to and then by the seed length. For each seed in the ranked list, from
the best to the worst, we drop the seed if it is already contained in an
alignment found before, or extend the seed with a banded affine-gap-penalty
dynamic programming (DP) if it potentially leads to a new alignment.

BWA-MEM's seed extension differs from the standard seed extension in two
aspects. Firstly, suppose at a certain extension step we come to reference
position $x$ with the best extension score achieved at query position $y$.
BWA-MEM will stop extension if the difference between the best extension score
and the score at $(x,y)$ is larger than $Z+|x-y|\times p_{\rm gapExt}$, where
$p_{\rm gapExt}$ is the gap extension penalty and $Z$ is an arbitrary cutoff.
This heuristic avoids extension through a poorly aligned region with good
flanking alignment. It is similar to the X-dropoff heuristic in
BLAST~\citep{Altschul:1997vn}, but does not penalize long gaps in one of the
reference or the query sequences.

Secondly, while extending a seed, BWA-MEM tries to keep track of the best
extension score reaching the end of the query sequence. If the difference
between the best score reaching the end and the best local alignment score is
below a threshold, the local alignment will be rejected even if it has a higher
score. BWA-MEM uses this strategy to automatically choose between local and
end-to-end alignments. It may align through true variations towards the end of
a read and thus reduces reference bias, while avoids introducing excessive
mismatches and gaps which may happen when we force a chimeric read into an
end-to-end alignment.

\begin{figure}[tb]
\centering
\includegraphics[width=0.4\textwidth]{alnroc-se}\\
\includegraphics[width=0.4\textwidth]{alnroc-pe}
\caption{Percent mapped reads as a function of the false alignment rate under
different mapping quality cutoff. $10^6$ pairs of 101bp reads are simulated
from the human reference genome using wgsim (http://bit.ly/wgsim2) with 1.5\%
substitution errors and 0.2\% indel variants. The insert size follows a normal
distribution $N(500,50^2)$. The reads are aligned back to the genome either as
single end (SE; top panel) or as paired end (PE; bottom panel). GEM is
configured to allow up to 5 gaps and to output suboptimal alignments (option
`--e5 --m5 --s1' for SE and `--e5 --m5 --s1 --pb' for PE). GEM does not compute
mapping quality. Its mapping quality is estimated with a BWA-like algorithm
with suboptimal alignments available. Other mappers are run essentially with
the default setting except for specifying the insert size distribution. The CPU
time in seconds is shown in the parantheses.}\label{fig:eval}
\end{figure}.

\subsection{Paired-end mapping}

\subsubsection{Rescuing missing hits}
Like BWA~\citep{Li:2009uq}, BWA-MEM processes a batch of reads at a time.
For each batch, it estimates the mean $\mu$ and the variance $\sigma^2$
of the insert size distribution from reliable single-end hits. For the
top 100 hits (by default) of each end, BWA-MEM performs SSE2-based
Smith-Waterman alignment (SW; \citealt{Farrar:2007hs}) for the mate within a
window $[\mu-4\sigma,\mu+4\sigma]$ from each hit. The second best SW score is
recorded to detect potential mismapping in a long tandem repeat. Hits found
from both the single-sequence alignment and SW rescuing will be used for
pairing. It is worth noting that BWA and BWA-SW only performs SW for the
best unique hit, but BWA-MEM also does SW for repetitive hits.

\subsubsection{Pairing} Given the $i$-th hit for the first read and $j$-th hit
for the second, BWA-MEM computes their distance $d_{ij}$ if the two hits are in
the right orientation, or sets $d_{ij}$ to infinity otherwise. It scores the
pair $(i,j)$ with $S_{ij}=S_i+S_j-\min\{-a\log_4 P(d_{ij}),U\}$, where $S_i$
and $S_j$ are the SW score of the two hits, respectively, $a$ is the matching
score and $P(d)$ gives the probability of observing an insert size larger than
$d$ assuming a normal distribution. `$\log_4$' arises when we interpret SW
score as odds ratio~\citep{Durbin:1998uq}. $U$ is a threshold that controls
pairing: if $d_{ij}$ is small enough such that $-a\log_4 P(d_{ij})<U$, BWA-MEM
prefers to pair the two ends; otherwise it prefers the unpaired alignments.
BWA-MEM chooses the pair $(i,j)$ that maximizes $S_{ij}$ as the final
alignments for both ends.  This pairing strategy jointly considers the
alignment scores, the insert size and the possibility of chimeric read pairs.

\end{methods}

\section{Results and Discussions}

We evaluated the performance of BWA-MEM on simulated data together with
NovoAlign (http://novocraft.com), GEM, Bowtie2, Cushaw2, BWA-SW and BWA
(Figure~\ref{fig:eval}). On accuracy, NovoAlign is the best. BWA-MEM is close
to NovoAlign for PE reads and is comparable to GEM and Cushaw2 for SE.  On
speed, BWA-MEM is similar to GEM and Bowtie2 for this data set, but is about
6 times as fast as Bowtie2 for a 650bp long-read data set.

To evaluate the performance for even longer query sequences, we aligned the
{\it E. coli} K-12 MG1655 substrain (AC:NC\_000913; 4.6Mb in length) against
the 536 strain (AC:NC\_008253) with both BWA-MEM and
nucmer~\citep{Kurtz:2004zr}. Nucmer finished the alignment in 25 seconds,
giving 105,505 substitution differences between the strains. BWA-MEM is slower.
It finished the alignment in 131 seconds, reporting 104,321 substitutions of
which 102,241 overlapping with the nucmer output. Manual examination of the
substitutions unique to each aligner reveals that most of them fall in short
regions with very high divergence. It is unclear which aligner is better. Note
that although nucmer is faster in the evaluation, only BWA-MEM scales well to
large genomes.

BWA-MEM is a fast and accurate aligner for sequence reads and is one of the few
that work well for both short reads and long sequences up to a few megabases.
Technically, it is possible to make BWA-MEM even faster for long sequences by
using SSE2-based banded DP and by restricting DP to regions not covered by a
long exact match. Seeding is the bottleneck for short sequences, while banded
DP is the bottleneck for long sequences.

\section{Acknowledgements}
\paragraph{Funding\textcolon} NIH 1U01HG005208-01.
\bibliography{bwamem}

\end{document}